---
title: >
  Report title
subtitle: >
  Bioinformatics Group, IMBEI, University Medical Center Mainz
author:
- name: Name Surname
  affiliation: 
  - Institute of Medical Biostatistics, Epidemiology and Informatics (IMBEI), Mainz, Germany
  email: yourmail@uni-mainz.de
## add here more authors if needed - and do add the affiliation(s)
date: "`r format(Sys.Date(), '%d %B %Y')`"
format:
  html:
    toc: true
    ## for clarity but also for cross-referencing
    number-sections: true
    ## this option makes the report one single file easiest to share
    embed-resources: true
    ## choosing a nice font
    mainfont: "Atkinson Hyperlegible"
    # mainfont: "IBM Plex Sans"
    ## making links automatically open in a new window 
    link-external-newwindow: true
    # code-link: true ## does seem to interfere with the title... pity.
    ## controls the aesthetics of the whole thing
    theme: 
    - cosmo
    ## if needed/desired: one can move that scss file and put it to toplevel
    - _extensions/bioinfo-html/ummz_theme.scss
editor_options: 
  chunk_output_type: console
## if needed/desired: one can move & rename that bib file and put it to toplevel
bibliography: _extensions/bioinfo-html/references.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  error    = FALSE,
  warning  = FALSE,
  eval     = TRUE,
  message  = FALSE
)
```

# My first section {#sec-firstsectionlabel}

The content of that section

# My second section {#sec-secondsectionlabel}

The text for that, and a reference inside the document to @sec-firstsectionlabel

Something looking like `code`.

## A subsection

In there, it is easy to refer to Anka's work [@Ludt2022]

# TO DO

Some code chunks and referring to their outputs nicely

```{r}
#| label: fig-myplot
#| fig-cap: "Plot caption"

plot(cars)
```

For example, see @fig-myplot.

```{r}
#| label: tbl-iris
#| tbl-cap: "Iris Data"

library(knitr)
kable(head(iris))
```


# An unnumbered section {-}

For example an appendix?

I can refer here to the previous @tbl-iris

# Session Information {-}

This document has been generated with...

```{r}
sessionInfo()
```

# References {-}
