---
title: "Using the BIUMmisc package for DE analyses"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BIUMmisc_DE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library("BIUMmisc")
```

**Compiled date**: `r Sys.Date()`

**Last edited**: `r Sys.Date()`

```{r loadLibraries, results= "hide"}
library("DESeq2")
library("topGO")
library("org.Hs.eg.db")
# library("pcaExplorer")
# library("ideal")
library("GeneTonic")
```

# Before running DE steps

1.	Load the expression data as `DESeqDataSet` object and create the associated annotation table.  
    Similar to Basic Protocol 3, load first the required packages, and create the fundamental `DESeqDataSet` object to be used for the analysis (using ENSEMBL identifiers) - optionally, one can filter the set of lowly expressed genes as specified in the chunk below. Generate the corresponding annotation table for `dds_macrophage`, and store that as `anno_df`.

```{r}
# Loading the packages
library("pcaExplorer")
library("ideal")
library("GeneTonic")

# Loading the data
library("macrophage")
library("DESeq2")
data("gse", package = "macrophage")
dds_macrophage <- DESeqDataSet(gse, design = ~line + condition)

# Changing the ids, removing the GENCODE-specific suffix
rownames(dds_macrophage) <- substr(rownames(dds_macrophage), 1, 15)
dds_macrophage
# Filtering low expressed features
keep <- rowSums(counts(dds_macrophage) >= 10) >= 6 
dds_macrophage <- dds_macrophage[keep, ]
dds_macrophage
```


```{r}
# Construct the annotation data frame
library("org.Hs.eg.db")
anno_df <- get_annotation_orgdb(dds = dds_macrophage,
                                orgdb_species = "org.Hs.eg.db",
                                idtype = "ENSEMBL")
```

```{r}
library("biomaRt")
mart <- useMart(biomart="ENSEMBL_MART_ENSEMBL", dataset="hsapiens_gene_ensembl")

anns <- getBM(attributes = c("ensembl_gene_id", "external_gene_name", "description"), 
              filters = "ensembl_gene_id",
              # filters = "external_gene_name",
              values = rownames(dds_macrophage), 
              mart = mart)
```


```{r}
plot_totcounts(dds_macrophage, group = "condition")
```

```{r}
extended_anno <- fortify_annotations(anno_df_1 = anno_df,
                                     anno_df_2 = anns,
                                     dds = dds_macrophage)

extended_anno$anno_df |> head()
extended_anno$dds

```


```{r}
tbl_normcounts <- deseq_normcounts_with_info(dds = dds_macrophage,
                                             extended_anno_df = extended_anno$anno_df)

tbl_TPMs <- deseq_tpm_with_info(dds = dds_macrophage,
                                extended_anno_df = extended_anno$anno_df)
```

# Running the DE steps

Creating the DE results themselves...


```{r}
dds_macrophage
design(dds_macrophage)
dds_macrophage <- DESeq(dds_macrophage)

resultsNames(dds_macrophage)

myresuSet_macrophage <- list()

myresuSet_macrophage <- 
  create_DEresults(resuSet = myresuSet_macrophage,
                   dds_obj = dds_macrophage,
                   contrast_name = "condition_IFNg_vs_naive", 
                   FDR = 0.05,
                   extended_anno_df = extended_anno$anno_df,
                   species = "Homo_sapiens"
                   )

myresuSet_macrophage <- 
  create_DEresults(resuSet = myresuSet_macrophage,
                   dds_obj = dds_macrophage,
                   contrast_name = "condition_IFNg_vs_naive", 
                   lfc_threshold = 0.6,
                   FDR = 0.05,
                   extended_anno_df = extended_anno$anno_df,
                   species = "Homo_sapiens",
                   name_res_entry = "ifng-naive-lfc0.6"
                   )

myresuSet_macrophage <- 
  create_DEresults(resuSet = myresuSet_macrophage,
                   dds_obj = dds_macrophage,
                   contrast_name = "condition_IFNg_vs_naive", 
                   lfc_threshold = 1,
                   FDR = 0.05,
                   extended_anno_df = extended_anno$anno_df,
                   species = "Homo_sapiens",
                   name_res_entry = "ifng-naive-lfc1"
                   )
```

```{r}
genes_of_interest <- c(
  "ENSG00000125347",   # IRF1
  "ENSG00000110944",   # IL23A
  "ENSG00000084636",   # COL16A1
  "ENSG00000172399"    # MYOZ2
)

plot_ma(res_obj = myresuSet_macrophage$`ifng-naive-lfc1`$res_DESeq, 
        intgenes = genes_of_interest)

plot_volcano(res_obj = myresuSet_macrophage$`ifng-naive-lfc1`$res_DESeq,
             ylim_up = 40,
             intgenes = genes_of_interest)

# TODO: adjust the y limit
# TODO: add text ready to be ggplotlified?
```

# After the DE steps: functional enrichment

```{r}
extended_anno_df <- extended_anno$anno_df
expressedInAssay <- (rowSums(assay(dds_macrophage))>0)
geneUniverseExprENS <- rownames(dds_macrophage)[expressedInAssay]
geneUniverseExpr <- extended_anno_df$gene_name[match(geneUniverseExprENS, extended_anno_df$gene_id)]
```

We iterate on all contrasts in the `myresuSet` object

```{r topgorun, results="hide", cache=TRUE}
for(i in names(myresuSet_macrophage)) {
  message(i)
  if(nrow(myresuSet_macrophage[[i]][["tbl_res_DE"]]) > 0) {
  myresuSet_macrophage[[i]][["topGO_tbl"]] <- 
    topGOtable(DEgenes = myresuSet_macrophage[[i]][["tbl_res_DE"]]$gene_name, 
               BGgenes = geneUniverseExpr, 
               ontology = "BP",
               geneID = "symbol",
               addGeneToTerms=TRUE, 
               topTablerows = 500,
               mapping = "org.Hs.eg.db")
  }
}

```


```{r cluprofload}
library("clusterProfiler")
```

```{r clupro, cache = TRUE}
for(i in names(myresuSet_macrophage)) {
  message(i)
  if(nrow(myresuSet_macrophage[[i]][["tbl_res_DE"]]) > 0) {
    myresuSet_macrophage[[i]][["clupro_tbl"]] <- 
      enrichGO(
        gene = myresuSet_macrophage[[i]][["tbl_res_DE"]]$gene_name,
        universe      = geneUniverseExpr,
        keyType       = "SYMBOL",
        OrgDb         = org.Hs.eg.db,
        ont           = "BP",
        pAdjustMethod = "BH",
        pvalueCutoff  = 0.01,
        qvalueCutoff  = 0.05,
        readable      = FALSE)
  }
}

```

# Wrapping up a DE report

transforming this into a gtl


```{r}
gtl_macs_ifng_naive <- resuset_to_gtl(myresuSet_macrophage, 
                                      result_name = "ifng-naive-lfc0.6",
                                      dds = dds_macrophage,
                                      anno_df = extended_anno_df)

describe_gtl(gtl_macs_ifng_naive)

vst_macrophage <- vst(dds_macrophage)
gs_heatmap(se = vst_macrophage,
           gtl = gtl_macs_ifng_naive,
           geneset_id = "GO:0034341",
           cluster_columns = TRUE,
           cluster_rows = TRUE,
           anno_col_info = "condition"
)

signature_volcano(gtl = gtl_macs_ifng_naive,
                  geneset_id = "GO:0034341",
                  FDR = 0.05
)

gs_scoresheat(
  mat = gs_scores(
    se = vst_macrophage,
    gtl = gtl_macs_ifng_naive),
  n_gs = 30
)
```


exporting them all

```{r}
exportr(myresuSet_macrophage,
        out_file_prefix = "fedetest", out_folder = "dirtest")
```



# Session information {-}

```{r}
BiocManager::version()
sessionInfo()
```

# References {-}


