# Samples exploration and assessment

We construct the dds objects which will contain all the required info for testing

```{r dds_rld, cache=TRUE}
counts <- fc_result$counts

dds <- DESeqDataSetFromMatrix(counts, samplesDesign, design = design_formula)
colnames(dds) <- samplesDesign$sampleID
rld_dds <- rlogTransformation(dds)
```

Remove all the rows where no signal could be detected at all:
```{r remove_rows}
dds <- dds[rowSums(counts(dds)) > 1, ]
nrow(dds)
```

The regularized log transformed values were used to perform sample clustering and principal component analysis (PCA).

```{r sampleDist}
pheatmap:::pheatmap(as.matrix(dist(t(assay(rld_dds)))))
```
[Comment on heatmap]



Create the PCA plots of all three datasets, again for comparison:
```{r pcaSamples}
top_genes <- c(500, 5000, 20000)
for (g in top_genes) {
  print(pcaExplorer:::pcaplot(rld_dds, ntop=g, title = paste0("top", g, " genes - global"), ellipse = F))
}
```
[Comment on PCA plot]


```{r Scree_plot}
pcaExplorer::pcascree(prcomp(t(assay(rld_dds[,])),center = T))
```

